---
title: "Financial_statement_analysis_with_Morningstar_data"
output: html_document
---
Export key ratios of Johnson & Johnson from Morningstar(http://financials.morningstar.com/ratios/r.html?t=JNJ)
Read the morning star data

```{r}
dat <- read.table(file="JNJ Key Ratios.csv", header=F, sep=',', 
                     dec = '.', stringsAsFactors=F, quote = "\"" , 
                     fill = TRUE , skip = 2)
```

The structure of the key ratio data. 

```{r}
str(dat)
```

Extract different sections of the financial data. 
This could apply to financial data of other companies on Morningstar. 
```{r}
financials <- dat[1:16, ]
profitability <- dat[c(18:27, 29:36),]
revenue <- dat[c(38, 40:48),]
net_income <- dat[c(38, 50:53),]
eps <- dat[c(38, 55:58),]
cash_flow <- dat[c(60:65),]
financial_health <- dat[c(67:92),]
efficiency_ratios <- dat[c(94:102),]
```

Clean up the profitability section. 
```{r}
# replace 1st row
profitability[1,] <- c("Date", 2005:2014, "TTM")
row.names(profitability) <- profitability[,1]
profitability <- profitability[,-1]
colnames(profitability) <- profitability[1,]
profitability <- profitability[-1,]
# remove the last column of TTM
profitability <- profitability[, 1:10]
```

Reorganize the data to make stacked area graph. 
```{r}

profitability[1:8,]
m <- matrix(0, nrow=50, ncol=3)
df <- data.frame(m)
colnames(df) <- c("Year", "OperatingMarginOrExpense", "PercentageOfSales")
# Force type to be character
df <- data.frame("Year"=numeric(50), "OperatingMarginOrExpense"=character(50), "PercentageOfSales"=numeric(50))

# add the first column "Year"
year <- sapply(2005:2014, function(x){rep(x, 5)})
year[1:length(year)]
df[,"Year"] <- year[1:length(year)]
# fill the 2nd column
df[,2] <- rep(c("COGS", "SG&A", "R&D", "Other", "Operating Margin"), 10)
df[,2]
# last column
temp <- vector()
PercentageOfSales <- sapply(1:10, function(x) {temp <- c(temp, profitability[c(2,4,5,6,7), x])})
df[,3] <- as.numeric(PercentageOfSales)
# Replace the NA with 0
df[is.na(df)] <- 0
```

Now, we can plot. 
First, the stacked area figure for different expenses. It shows the relative 
proportion of those expenses. 
```{r}
library(ggplot2)
ggplot(df, aes(x=Year, y=PercentageOfSales, fill=OperatingMarginOrExpense)) + 
        geom_area(colour="black", size=0.2, alpha=0.5) + 
        scale_x_continuous(breaks= c(2005:2014), limits=c(2005, 2014)) + 
        xlab("Year") + ylab("% of Sales")
```


Sales
```{r}
# do sales
# remove last column of financials
financials <- financials[,1:10]
financials
# create a df2
df2 <- matrix(0, nrow=10, ncol=2)
df2 <- data.frame(df2)
# fill in df2 with sales data
df2[,2] <- as.character(financials[1,])
df2[,1] <- colnames(financials)
df2
# add colnames
colnames(df2) <- c("Year", "Sales")
df2
# add group=1 for factor type
ggplot(df2, aes(x=Year, y=Sales, group=1)) + geom_line()
```

```{r}
# plot stacked bar figure of operating cash flow, including cap spending 
# free cash flow
# data already cleaned as above

# extract data for the plot

df3 <- financials[12:13,]

# add a row with year
df3 <- rbind(as.numeric(colnames(df3)), df3)
# rename rows
rownames(df3) <- c("Year", "Capital_Spending", "Free_Cash_Flow")

# Transpose
df3 <- t(df3)

# transform data type to numeric first
(df3 <- gsub(",", "", df3))
# negateive to positive
df3 <- gsub("-", "", df3)
# transform into data frame
df3 <- data.frame(df3)
# remove the last row
df3 <- df3[1:10,]
# melt the data frame
library(reshape)
?melt
df4 <- melt(df3, id.vars="Year", measure.vars=c("Capital_Spending", "Free_Cash_Flow"), variable_name="Type")

# rename col3 into 
colnames(df4)[3] <- "Operating_Cash_Flow"
# change to numeric, to avoid discrete y ticks
df4$Operating_Cash_Flow <- as.numeric(as.character(df4[,3]))

# let's do the stacked bar plot
ggplot(df4, aes(y=Operating_Cash_Flow, x=Year, fill=Type)) +
        geom_bar(stat="identity") +
        guides(fill=guide_legend(reverse=TRUE)) + 
        ylab("Operating Cash Flow (Millions)")
        #theme(axis.text.y=element_blank()) +
        #scale_y_discrete(limits=)
```

Liquidation ratios
```{r}
# liquidity ratios
# in financial_health

df5 <- financial_health[22:26,]        
df5 <- as.data.frame(t(df5))
colnames(df5) <- c("Year", "CurrentRatio", "QuickRatio", "FinancialLeverage", "DebtToEquity")
# remove first row
df5 <- df5[-1,]

# melt df5 for ggplot
df6 <- melt(df5, id.vars="Year", measure.vars=c("CurrentRatio", "QuickRatio", "FinancialLeverage", "DebtToEquity"), 
            variable_names="Ratios")
# change to numeric, to avoid discrete y ticks
df6$value <- as.numeric(as.character(df6$value))

# multiple bar plot
ggplot(df6, aes(x=Year, y=value, fill=variable)) + 
        geom_bar(position="dodge", stat="identity") +
        theme(axis.text.x = element_text(angle=30))
        #theme(axis.title.x=element_text(angle=45))
```


The whole process could be automated if we export data directly from the website, without downloading the .csv file locally. 