---
title: "Baloon_plot_with_GSEA_result"
output: html_document
---

Read the data for baloon plot. The data was exported after gene set enrichment analysis (http://software.broadinstitute.org/gsea/index.jsp). There are three variables: NAME (names of gene sets), NES (normalized enrichment score), and sig_genes (number of genes significantly changes in corresponding gene sets).  

```{r}
dt <- read.csv("data_for_baloon_plot.csv")
head(dt)
```

Plot to have a look
```{r}

library(ggplot2)
p <- ggplot(dt, aes(x=NES, y=NAME, size=sig_genes, 
                        colour=1, fill=1)) + 
        geom_point(shape=21) +
        xlab("Normalized enrichment score") + ylab("Selected hallmark gene sets")
p
```

Color baloons according to p values
Mark red if p value < 0.1
```{r}
dt2 <- read.csv("data_for_baloon_plot_p_val.csv")
# Add color info
col <- vector()
for (i in 1:length(gsea2[,1])) {
        if (gsea2$NOM.p.val[i] <= 0.1) {
                col[i] <- 1
        }
        else {col[i] <- 3}
}
# Add col column to bub_dt2
dt2$color <- col
# remove "HALLMARK_" in names
dt2$NAME <- gsub("^.*?_","",bub_dt2[,1])

# plot again
p <- ggplot(dt2, aes(x=NES, y=NAME, size=sig_genes, colour=color, fill=color)) + 
        geom_point(shape=21) + 
        xlab("Normalized enrichment score") + ylab("Selected hallmark gene sets")

p
```

Details to improve next. Color and color legends. 
